---
title: "SQL"
output: html_document
---
## Filtering Rows and Colums

In SQL you can filter columns by using SELECT and rows by using WHERE.Let's look at an example using the Lahman baseball database. First we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

### Example:
Suppose we want to see the homerun totals for the 1927 Yankees. 
We could write:
```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA'and yearID=1927"

sqldf(query)
```
# Baseball Example
## SQL

If one wants to find the specific codes of words for Batting they would use 'colnames(Batting)'. In these examples, SQL is being used. First one needs to load these packages:

```{r}
library(Lahman)
library(sqldf)
```

### Example: 
Find all instances where the Yankees have hit fourty or more homeruns.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and HR>=40"

sqldf(query)
```

### Example:
Find all instances where a Yankee had more than fourty homeruns but less than sixty strikeouts.

```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"

sqldf(query)
```

### Example:
Find all instances of Phillies hitting more than thirty homeruns in the 1970s.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"

sqldf(query)
```

## ORDER BY

With SQL, we can also order the rows in the output.
### Example:
suppose you wanted to find every instance of a player hitting more than fifty homeruns. Suppose you want to have the players with the highest homeruns at the top, you can do this the following way:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50 ORDER BY HR DESC"
sqldf(query)
```

### Example:
Find all instance of a player striking out less than 10 times, make sure each player has at least 400 at-bats (AB)

```{r}
query<"SELECT playerID,teamID,yearID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
order by SO"
sqldf(query)
```
## AGGREGATION

```{r}
query<-"SELECT playerID,HR FROM Batting 
WHERE playerID='ruthba01'"
sqldf(query)
```

### Example:

Find the career homerun totals for Babe Ruth.
```{r}
query<-"SELECT playerID,sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

### Example:

find the career homerun totals for all players, but list only those with a total of 600 or mor. order by having greatest total at the top.
```{r}
query<-"SELECT playerID,sum(HR) FROM Batting 
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"
sqldf(query)
```

### Example:

What players have the highest average per season homerun total.
```{r}
query<-"SELECT playerID,avg (HR) FROM Batting
GROUP BY playerID
HAVING avg(HR) >30
ORDER BY avg(HR) DESC"
sqldf(query)
```

## Joins

When one wants to join two tables together one must specify with the use of 'inner join' when specifying from. In addition one must specify which variables are shared between the two tables.

### Example:
Join Batting table with Master table to display nameFirst and nameLast instead of playerID as 'ruthba01'.
```{r}
query<-"select nameFirst, nameLast, teamID, yearID,HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where Batting.playerID = 'ruthba01'"

sqldf(query)
```
### Example:
Join Batting and Master to display names of players with HR greater than 50 in descending order.
```{r}
query<-"select nameFirst, nameLast, teamID, yearID, HR 
from Batting inner join Master
on Batting.playerID=Master.playerID
where HR>50
order by HR desc"

sqldf(query)
```

### Example:
Join Batting and Teams tables as to display the full name of team instead of teamID for the player 'ruthba01'
```{r}
query<-"select playerID, name, Batting.yearID, Batting.HR
from Batting inner join Teams
on Batting.teamID=Teams.teamID
and Batting.yearID=Teams.yearID
where playerID='ruthba01'"

sqldf(query)
```

### Example:
Joining of three tables together(Batting, Teams, Master) and identifying 'ruthba01'. When joining three tables use '()' when first join is written.
```{r}
query<-"select nameFirst, nameLast, name, Batting.yearID, Batting.HR 
from (Batting inner join Master 
on Batting.playerID=Master.playerID) inner join Teams 
on Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
where Batting.playerID='ruthba01'"

sqldf(query)
```
### Example:
Join Batting and Master to display players that have a career total of more than 600 HR.
```{r}
query<-"select nameFirst, nameLast, sum(HR)
from Batting inner join Master 
on Batting.playerID = Master.playerID
group by Batting.playerID
having sum(HR)>600
order by sum(HR) DESC"

sqldf(query)
```

### Example:
List all players with the first name Bob, who averaged more than one million dollars in salary over their career. Include last names. You will need to use the colnames(Salaries) table.
```{r}
query<-"SELECT nameFirst,nameLast,salary FROM Salaries INNER JOIN Master On Salaries.playerID=Master.playerID GROUP BY Master.playerID HAVING nameFirst='Bob' and salary>=1000000"
sqldf(query)
```






